---
title: "WorkStateCategory"
output: html_document
date: "2026-02-16"
---
``` {r}
library(readr)
StateAndCategory <- read_csv("StateAndCategory.csv")
```
``` {r}
library(tidyverse)
StateAndCategory$...14 <- NULL
StateAndCategory$...15 <- NULL
StateAndCategory$...16 <- NULL
StateAndCategory$...17 <- NULL
StateAndCategory$...18 <- NULL
```

```{r}
States <- StateAndCategory[1:88924, ]
```

```{r}
s <- gsub("$","", States$Dollars, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$Dollars <- as.numeric(s)
```

```{r}
s <- gsub("$","", States$`Dollars last year`, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$`Dollars last year` <- as.numeric(s)
```

```{r}
s <- gsub("$","", States$`Dollars 3 years ago`, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$`Dollars 3 years ago` <- as.numeric(s)
```

```{r}
library(ggplot2)

ggplot(States, aes(`Unit sales`)) + geom_density(alpha = 0.2)

```

```{r}
library(ggplot2)
ggplot(States,
       aes(x = as.Date(Date), y = `Unit sales`, color = Category)) +
  geom_line(alpha = 0.7) +
  theme_minimal() +
  labs(title = "Unit Sales Over Time",
       x = "Date",
       y = "Unit Sales",
       color = "Category")
```

```{r}
States$`Unit price` <- States$Dollars/States$`Unit sales`
```

```{r}
library(readr)
write_csv(States, "outputstates.csv")
```
to import data, we decided to remove 2019 and 2023 from the population data set so that we can more simply average weeks of groceries to make an interactive map per capita
claude sonnet 4.6 extended on 3/11/26 sources data for state population
claude sonnet 4.6 made new data table on 3/11/26 to average unit prices and dollars spent per year per state

```{r}
library(readr)
library(tidyverse)
us_state_populations_2019_2023 <- read_csv("us_state_populations_2019_2023.csv")
us_state_populations_2019_2023[6] <- NULL   # remove 6th column first!
us_state_populations_2019_2023[2] <- NULL   # then 2nd column
us_state_populations_2019_2023 <- us_state_populations_2019_2023[-c(2, 8, 12, 13, 27, 31, 35), ]
us_state_populations_2020_2022 <- us_state_populations_2019_2023
View(us_state_populations_2020_2022)
summary(us_state_populations_2020_2022)
```

``` {r}
library(tidyr)
library(dplyr)
df <- read.csv("avg_spending_by_state_year_category.csv", check.names = FALSE)
df_wide <- df %>%
  select(State, Year, Category, Avg_Dollars_Spent) %>%
  spread(key = Category, value = Avg_Dollars_Spent)
 
print(df_wide)
```

```{r}
library(dplyr)
library(tidyr)
library(readr)

# Read the data
data <- read_csv("avg_spending_by_state_year_category.csv")

# Get categories
categories <- unique(data$Category)

all_tables <- list()

for(cat in categories){

  category_table <- data %>%
    filter(Category == cat) %>%
    group_by(State, Year) %>%
    summarize(avg_spending = mean(Avg_Dollars_Spent, na.rm = TRUE)) %>%
    spread(key = Year, value = avg_spending)

  # Convert wide table back to long format
  category_long <- category_table %>%
    gather(key = Year, value = Avg_Dollars_Spent, -State) %>%
    mutate(Category = cat)

  all_tables[[cat]] <- category_long
}

# Combine all categories
final_data <- bind_rows(all_tables)

# Reorder columns
final_data <- final_data %>%
  select(State, Year, Category, Avg_Dollars_Spent)

# View result
final_data

# Save file
write_csv(final_data, "state_year_category_spending.csv")
```
claude sonnet 4.6 made new data table on 3/16/26 to create new tables for each category that give average dollars spent and state and year. We do not need them for aything as of 3/16 but are keeping code

```{r}
population <- read.csv("us_state_populations_2020_2022") %>%
  gather(key = "Year", value = "Population", "2020", "2021", "2022")

```
