---
title: "WorkStateCategory"
output: html_document
date: "2026-02-16"
---
``` {r}
library(readr)
StateAndCategory <- read_csv("StateAndCategory.csv")
```
``` {r}
library(tidyverse)
StateAndCategory$...14 <- NULL
StateAndCategory$...15 <- NULL
StateAndCategory$...16 <- NULL
StateAndCategory$...17 <- NULL
StateAndCategory$...18 <- NULL
```

```{r}
States <- StateAndCategory[1:88924, ]
```

```{r}
s <- gsub("$","", States$Dollars, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$Dollars <- as.numeric(s)
```

```{r}
s <- gsub("$","", States$`Dollars last year`, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$`Dollars last year` <- as.numeric(s)
```

```{r}
s <- gsub("$","", States$`Dollars 3 years ago`, fixed = TRUE)
s <- gsub(",","", s, fixed = TRUE)
States$`Dollars 3 years ago` <- as.numeric(s)
```

```{r}
library(ggplot2)

ggplot(States, aes(`Unit sales`)) + geom_density(alpha = 0.2)

```

```{r}
library(ggplot2)
ggplot(States,
       aes(x = as.Date(Date), y = `Unit sales`, color = Category)) +
  geom_line(alpha = 0.7) +
  theme_minimal() +
  labs(title = "Unit Sales Over Time",
       x = "Date",
       y = "Unit Sales",
       color = "Category")
```

```{r}
States$`Unit price` <- States$Dollars/States$`Unit sales`
```

```{r}
library(readr)
write_csv(States, "outputstates.csv")
```
to import data, we decided to remove 2019 and 2023 from the population data set so that we can more simply average weeks of groceries to make an interactive map per capita
claude sonnet 4.6 extended on 3/11/26 sources data for state population
claude sonnet 4.6 made new data table on 3/11/26 to average unit prices and dollars spent per year per state

```{r}
<<<<<<< HEAD
library(readr)
library(tidyverse)
us_state_populations_2019_2023 <- read_csv("us_state_populations_2019_2023.csv")
us_state_populations_2019_2023[6] <- NULL   # remove 6th column first!
us_state_populations_2019_2023[2] <- NULL   # then 2nd column
us_state_populations_2019_2023 <- us_state_populations_2019_2023[-c(2, 8, 12, 13, 27, 31, 35), ]
us_state_populations_2020_2022 <- us_state_populations_2019_2023
View(us_state_populations_2020_2022)
summary(us_state_populations_2020_2022)
```
=======
library(tidyr)
library(dplyr)
df <- read.csv("avg_spending_by_state_year_category.csv", check.names = FALSE)
df_wide <- df %>%
  select(State, Year, Category, Avg_Dollars_Spent) %>%
  spread(key = Category, value = Avg_Dollars_Spent)

print(df_wide)
```
>>>>>>> 24949781e7e4f036ab5539010d1c106b97f5b4ee
